### 2023-12-11 docs/latex/paper.tex
Finished Method's Section

\documentclass{article}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables
\usepackage{hyperref}
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\usepackage[accepted]{mlsys2024}

\mlsystitlerunning{Step-by-Step: Training IMU-based Gestures with Live Feedback}

\begin{document}
\twocolumn[
\mlsystitle{Step-by-Step: Training IMU-based Gestures with Live Feedback}

\begin{mlsysauthorlist}
\mlsysauthor{Michael Schnebly}{h}
\end{mlsysauthorlist}

\mlsysaffiliation{h}{School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, USA}
\mlsyscorrespondingauthor{Michael Schnebly}{michael\_schnebly@g.harvard.edu}

\mlsyskeywords{Machine Learning, MLSys}

\vskip 0.3in

\begin{abstract}
Recognizing user-defined gestures in inertial measurement unit (IMU) data unlocks new forms of creativity and accessibility in human-computer interaction. However, training gesture recognition models is a difficult task that requires a deep understanding of machine learning. We present Step-by-Step, a software tool that allows users to train gesture recognition models with live audiovisual feedback. Step-by-Step uses a simple neural network to learn to recognize and distinguish multiple gestures in IMU timeseries data. Users can train the model by performing gestures and receiving live feedback on the model's performance. Step-by-Step is designed to be accessible to users with no machine learning experience, while providing a powerful codebase for advanced users.
\end{abstract}
]

\printAffiliationsAndNotice{}

\section{Introduction}

\section{Background}

\section{Method}
% Introductory paragraph summarizing the method's components: hardware, data collection, data labelling, model training, model inference, user interface, and the software architecture.
% Hardware: what hardware is used, and how it is connected to the computer.
% Data collection: how the data is collected, what data is collected, and how it is stored.
% Data labelling: how the data is labelled, what the labels are, and how they are stored.
% Model structure: what the model is, how it is structured, and how it is stored.
% Model training: how the model is trained, what the model is, and how it is stored.
% Model inference: how the model is used to make predictions, and how the predictions are stored.
% User interface: how the user interacts with the system, and how the system interacts with the user.
% Software architecture: how the system is structured, and how the components interact with each other.

\subsection{Hardware}
For measuring body movement, we used the Bosch BNO-055, a 9-axis IMU that includes an accelerometer, gyroscope, and magnetometer. The IMU uses I2C communication protocol to stream data to a microcontroller (Espressif ESP8266) which passes that data on to a laptop  (Macbook Pro 2017) via USB for further processing.

Note that the particular hardware components used in this project are not essential to the system. They could be replaced with other options so long as the sensor data contains information necessary to recognize and distinguish the gestures of interest and the computer (or microcontroller if using purely embedded approach) is capable of running the software at a rate matching that of data collection.
% The hardware is shown in Figure \ref{fig:hardware}.

\subsection{Software}
The software is written in Python and designed to be accessible to users with no machine learning experience while providing a powerful codebase for advanced users. The software is structured into three main components: data processing, neural network, and user interface. The data processing component handles data collection, labelling, and storage. The neural network component handles model structure, training, and inference. The user interface component handles user interaction and feedback.

To facilitate realtime training and inference, the software is structured as a pipeline with the following processes running in parallel: data processing, model training, and model inference. Importantly all processes refer to the same global state, allowing the training data and model to be updated and tested in realtime as new data is collected.
% The software architecture is shown in Figure \ref{fig:software_architecture}.

\subsection{Data Processing}
\subsubsection{Collection}
The IMU data is collected as a timeseries of frames. Each frame is a 3D vector representing linear acceleration in each of the three axes: x, y, and z. The data is collected at a rate of 100 Hz and can be streamed directly from an IMU, recorded to a file, or loaded from a file.

\subsubsection{Labelling}
The data is labelled by associating each frame with a gesture. A gesture is a sequence of frames that represents a single movement of the body. Here, we focus on percussive gestures and assume that a gesture is complete at the local peak in linear acceleration's magnitude. This constraint allow us to achieve realtime, automated labelling of gestures using a simple peak-detection algorithm. When a local peak meet's user-defined criteria (e.g. acceleration is above a certain threshold), the peak frame is labelled as containing a gesture. All other frames are labelled as containing no gesture.

\subsection{Neural Network}
\subsubsection{Structure}
The neural network is a shallow, two-branched model. One branch represents a recent history of sensor values while the other represents a recent history of model outputs. Combining the information from the these two sources, the model is trained to maximize the probability of the correct gesture and minimize the probabilities of all other gestures.

The sensor branch consists of an input layer (short timeseries of recent sensor values) followed by a  a 1D convolutional layer (recognizes patterns in the timeseries). The memory branch consists of an input layer (short timeseries of recent model outputs) followed by a max pooling layer (summarizes the timeseries of recent model outputs based on its highest values). These branches are then concatenated and followed by a fully connected output layer. 
% The model structure is shown in Figure \ref{fig:model_structure}.

\subsubsection{Realtime Training}
Training is performed using Tensorflow's default gradient descent optimizer. The training data consists of a timeseries of frames, each of which is labelled with a gesture. The model is trained on a sliding window of frames where window size is a user-defined parameter. For example, window size could be set to 20 frames, equating to 200ms of data at 100Hz.

The window slides forward one frame at a time. If the most recent frame in a window is labelled with a gesture, the window is labelled with that gesture. Otherwise, the window is labelled with no gesture. As new windows are collected, the model is trained on the windows using gradient descent. Training is iterative and the model's training data is updated in realtime, allowing the user to receive live feedback on the model's performance.

In principle this training could be done stochastically, training on each frame's window as it is collected. However, this would  would require a more sophisticated training algorithm that Tensorflow's built-in gradient descent optimizer (at least given the computational performance of the author's machine). For this reason, this proof-of-concept codebase defaults to training on batches of recent windows. When training completes, the model is updated, the training data is updated with recent frames, and training begins again.

\subsubsection{Realtime Inference}
As the model is updated, its performance is tested on the most recent frames using Tensorflow. Along with the groundtruth labels, the model's predictions are used to generate live feedback for the user using both auditory and visual cues as discussed in the User Interface section.

\subsection{User Interface}
The user interface is designed to provide live feedback on the model's performance. The user interface consists of two components: a visual component and an auditory component. The visual component is a simple GUI that displays the model's predictions and groundtruth labels in realtime. The auditory component is a simple synthesizer that plays a tone corresponding to the model's prediction.

Using keyboard inputs, the user can control various aspects of the application. The user can start and stop data collection, start and stop model training, and start and stop model inference.
% The user interface is shown in Figure \ref{fig:user_interface}.

\section{Findings}

\bibliography{example_paper}
\bibliographystyle{mlsys2024}

\end{document}